\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-04-25 arch.tex
updates to review

\begin{figure}[t]
  \centering
  \subfloat[\small A \rx agent is composed of multiple reactors or
    control loops (rounded boxes) which are connected through state
    variables provided by one reactor ({\color{blue}blue} solid line)
    with multiple possible clients ({\color{green}green} dashed
    lines)]{\includegraphics[scale=0.45]{figs/AUV-agent.pdf}
  \label{fig:agent}
  }
  \qquad
  \subfloat[\small State information flow between the \texttt{Pilot}
  and the \texttt{Vehicle} reactors in an \rx instance.  Observations
  transit as a bottom-up flow and apply to the execution frontier
  $\tau$.  Goal information flow from the top to the bottom and
  relates to desired future values of state variables.]{
    \includegraphics[width=0.4\columnwidth]{figs/TREX_flow}
    \label{fig:trex:flow}
  }
\caption{\small \rx reactors and shared state variables.}
\end{figure}

@@

\item {\em Observation on current state values}: This information
  propagates \emph{up} in the reactor dependency graph at the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variables for this tick so they can
  compute their {\em internal} state variable values for this same
  tick. By doing so we propagate a consistent view of that state of
  the world throughout all the reactors as time
  advances.

@@

\subsubsection{Reactors failure handling in \rx}
\label{sec:rx-reactor-failure}
It is important to note that, in T-REX architecture, the only critical
task of a reactor is to successfully identify its state through
synchronization as the execution frontier advances. Indeed, a reactor
that fails to integrate its {\em external} state updates and/or
identify its {\em internal} state at any tick express that it is not
able anymore to explain the state of the world. In the agent point of
view that means that any state information from this reactor is not to
be trusted anymore and results on this reactor exclusion from the
execution cycle and consequent destruction. As the agent destroy 
the  faulty reactor is destroyed the agent produced on all of 
its former {\em internal} timelines the special observation 
\texttt{Failed} to all the reactors which did declare these state 
variables as {\em external} informing
them through this that this specific state variable is currently
unavailable. On the other hand all goals exchanged between reactors
are considered as rejectable by the reactor that manages them. As a
result the failure to find a plan for these goals, while to be avoided,
is not considered as a critical failure of the reactor and will only
result on these goals being ignored until new goals are received.

Take for example our AUV agent from Fig. \ref{fig:agent},
and assume that the \texttt{Sampling} reactor failed to synchronize;
in  such case this reactor is destroyed and both
\texttt{Sampling\_policy} and \texttt{Relevant\_data} state variables
produced the \texttt{Failed} observation. Therefore, the
\texttt{Survey controller} will received this observation on its {\em
  external} \texttt{Sampling\_policy} which indicates that this state
variable is not managed anymore. At this point depending on this
reactor implementation we have two possible outcomes:
\begin{enumerate}
\item \texttt{Survey  controller} model allows him to integrate this
  \texttt{Failed} observation. For example it can identify that, from
  now on, there's no way to take water samples. It still can 
  navigate as \texttt{Pilot} is available which allow him to continue
  this aspect of its current plan.
\item \texttt{Survey controller} models is not able to handle this
  unexpected failure resulting in turn on its failure to
  synchronize. This result on this reactor being also destroyed and
  the \texttt{Failed} observation is produced on the \texttt{survey}
  timeline.
\end{enumerate}
Even on the worst case scenario presented on this second case, we
note that a reactor failure only propagates up in the reactor graph. 
In our example, even if all the reactors that depends on
\texttt{Sampling} are entrained by its failure, we still keep the
three reactors \texttt{Vehicle}, \texttt{Pilot} and \texttt{Ship
  operation} allowing to  maintain the vehicle navigation control
(\texttt{Pilot}) along with high level safety operation 
(\texttt{Ship operation}). It illustrates how \rx can handle graceful
system degradation despite the failure of one or more reactors. It
also emphasize that it is critical to ensure that the low level
reactors are properly designed in order to be as robust as 
possible.

>>>>

\begin{figure}[t]
  \centering
  \subfloat[\small A \rx agent is composed of multiple reactors or
    control loops (rounded boxes) which are connected through state
    variables provided by one reactor ({\color{blue}blue} solid line)
    with multiple possible clients ({\color{green}green} dashed
    lines).]{\includegraphics[scale=0.45]{figs/AUV-agent.pdf}
  \label{fig:agent}
  }
  \qquad
  \subfloat[\small State information flow between the \texttt{Pilot}
  and the \texttt{Vehicle} reactors in an \rx instance.  Observations
  transit as a bottom-up flow and apply to the execution frontier
  $\tau$.  Goal information flow from the top to the bottom and
  relates to desired future values of state variables.]{
    \includegraphics[width=0.4\columnwidth]{figs/TREX_flow}
    \label{fig:trex:flow}
  }
\caption{\small \rx reactors and shared state variables.}
\end{figure}

@@

\item {\em Observation on current state values}: This information
  propagates \emph{up} in the reactor dependency graph at the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variables for this tick so they can
  compute their {\em internal} state variable values for this same
  tick. By doing so we propagate a consistent view of that state of
  the world throughout all the reactors as time
  advances. \kcomment{Observations can therefore be both sensor
    measurements as well state passed up from other reactors. The
    strict rules of timeline ownership enable a clear policy for
    conflict resolution: observations dominate expectations.}

@@

\subsubsection{Handling Failure in \rx}
\label{sec:rx-reactor-failure}

The critical task of a \rx reactor is to successfully identify its
state through synchronization as the execution frontier advances. When
a reactor fails to integrate its {\em external} state update and/or
identify its {\em internal} state at any tick, it implies that it is
unable to explain the state of the world.  This in turn implies that
any state information from this reactor is suspect, resulting in the
exclusion of this reactor from the execution cycle and consequent
destruction. As the agent destroys the faulty reactor, a special
observation, \texttt{Failed}, is produced on all {\em internal}
timelines. Reactors which declare these very state variables as {\em
  external}, are informed of the failure of this timeline.  All goals
exchanged between reactors are considered \emph{rejectable} by the
reactor that manages them. As a result in such a situation, the
failure to find a plan for these goals, is \emph{not} considered a
critical failure of the reactor. It will only result \kcomment{in}
these goals being ignored until newer goals are received.

Take for example our AUV agent from Fig. \ref{fig:agent}, and assume
that the \texttt{Sampling} reactor failed to synchronize. In such a
case this reactor is destroyed and both \texttt{Sampling\_policy} and
\texttt{Relevant\_data} timelines generate the \texttt{Failed}
observation as illustrated in Fig. \ref{fig:fail0}.  The
\texttt{Survey controller} will receive this \texttt{Failed}
observation on its {\em external} \texttt{Sampling\_policy} timeline
which indicates that this state variable is no longer managed by its
owner reactor. This leads to two potential outcomes based on how the
reactor is implemented:

\item The \texttt{Survey controller} model allows it to integrate the
  \texttt{Failed} observation. For example it can identify that,
  subsequent use of the water sampler is no longer possible.  The \rx
  agent can continue to navigate since \texttt{Pilot} is still
  available. In such a case the agent structure remains as in
  Fig. \ref{fig:fail0} with no consequent impact.

\item The \texttt{Survey controller} model is unable to handle this
  unexpected failure resulting in its failure to synchronize. This
  results in the \texttt{Survey controller} also failing, with
  subsequent destruction and flagging via the \texttt{Failed}
  observation on its \texttt{survey} timeline as shown in
  Fig. \ref{fig:fail1}.

\begin{figure}[b]
  \centering
  \subfloat[\small When the \texttt{Sampling} reactor fails to
  synchronize, it is removed from the reactor dependency graph along
  with all its {\em internal} timelines while generating a \texttt{Failed} observation]{\includegraphics[width=0.35\columnwidth]{figs/AUV-fail0.pdf}
  \label{fig:fail0}}
  \qquad
  \subfloat[\small In the worst case this \texttt{Failed} observation may imply
  other reactors that depend on the \texttt{Sampling} reactor might fail to
  synchronize leading to a bottom-up destruction of ``faulty'' reactors.]{
    \includegraphics[width=0.35\columnwidth]{figs/AUV-fail1}\label{fig:fail1}}
  \caption{\small Impact of the \texttt{Sampling} reactor failure in
    the AUV agent \kcomment{from} Fig. \ref{fig:agent}}
\end{figure}

Even in the worst of the two cases presented above, we note that a
reactor failure only propagates \emph{up} in the reactor dependency
graph.  In our example, even if all the reactors that depend on
\texttt{Sampling} are entrained by its failure, \rx continues to keep
the reactors \texttt{Vehicle}, \texttt{Pilot} and \texttt{Ship
  operation} intact.  This ensures maintenance of vehicle control via
the \texttt{Pilot}, in addition to an abstract commanding mode for
safe operation via the \texttt{Ship operation} reactor, illustrating
how \rx can handle system degradation despite the failure of one or
more reactors, and doing so gracefully.

Consequently, by ensuring that lower level reactors closest to the
platform hardware are robustly designed, the \rx framework ensures
that system designers can focus their attention on critical components
while factoring in deliberative control of their platforms.

### 2012-04-25 concepts.tex
updates to review

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.35]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\end{figure}

@@

\eu uses a \emph{state variable} representation
\cite{mus94,jonsson00,mcgann09} to describe the evolution of state
over time. The instantiated history of such state variable evolution
over a temporal horizon we call \emph{timelines} and which represent a
single thread in the execution of a concurrent system. At any given
time each thread can execute a single procedure.  Thus each timeline
consists of a sequence of procedures which encapsulate and describe
state evolution; we call these instantiated atomic entities
\emph{tokens}.  A token therefore describes a procedure invocation,
the state variables on which it can occur, the parameter values of the
procedure, and the time values defining the interval. We allow
encapsulation of uncertainty within these tokens with a range of start
and end times and parameters, all of which are encoded as variables. A
constraint solver in turn manipulates these variables defined in a \eu
domain model. For example, consider a scientific need to take water
samples 100 meters from an environmental hotspot while an AUV is
performing a Yo-Yo sawtooth pattern in the water-column. Two samples
are needed if the feature's signal is above a threshold; one
otherwise. The token that is capturing the sensory threshold has a
parametric constraints to the token which fires the requisite water
sampler. In addition, the start time of the water sampling procedure
is highly dependant on the variability of sub-sea currents and actual
speed of the vehicle. Therefore a number of values are possible for
the start times and duration of the sampling all of which are valid
combinations for desired outcomes.

>>>>

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.4]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\end{figure}

@@

\eu uses a \emph{state variable} representation
\cite{mus94,jonsson00,mcgann09} to describe the evolution of state
over time. The instantiated history of such state variable evolution
over a temporal horizon we call \emph{timelines} and which represent a
single thread in the execution of a concurrent system. At any given
time each thread can execute a single procedure.  Thus each timeline
consists of a sequence of procedures which encapsulate and describe
state evolution; we call these instantiated atomic entities
\emph{tokens}.  A token therefore describes a procedure invocation,
the state variables on which it can occur, the parameter values of the
procedure, and the time values defining the interval. We allow
encapsulation of uncertainty within these tokens with a range of start
and end times and parameters, all of which are encoded as variables. A
constraint solver in turn manipulates these variables defined in a \eu
domain model. For example, consider a scientific need to take water
samples 100 meters from an environmental hotspot while an AUV is
performing a Yo-Yo sawtooth pattern\footnote{\kcomment{Variability in
    the water column is along the vertical dimension. Since the Dorado
    platform can only move forward, the Yo-Yo pattern is the most
    efficient mechanism for studying water column properties.}}  in
the water-column. Two samples are needed if the feature's signal is
above a threshold; one otherwise. The token that is capturing the
sensory threshold has a parametric constraints to the token which
fires the requisite water sampler. In addition, the start time of the
water sampling procedure is highly dependant on the variability of
sub-sea currents and actual speed of the vehicle. Therefore a number
of values are possible for the start times and duration of the
sampling all of which are valid combinations for desired outcomes.
